---
title: "Proportions Test (MSM vs IDU)"
date: "October 21, 2015"
output: html_document
---
###**Proportions Test Analysis**
####*Comparing deaths of Men who have Sex with Men (MSM) and Injected Drugs Users (2003-2014)*
<br>
In order to conduct this analysis we used data from the HIV Surveillance System for 2003 to 2014. You can access the raw file [here.](https://github.com/kramus/proptest/blob/master/propt3.csv)
<br>  
The structure of the data is as follows:
```{r, include=FALSE}
library(plyr)
library(dplyr)
library(lattice)
library(ggplot2)
library(markdown)
library(pander)
library(knitr)
library(xtable)
```
```{r, echo=FALSE}
data <- read.csv(".\\propt3.csv")
data$case <- factor(data$case, levels(data$case)[c(2,1)]) 
#one way of changing order of levels
data$type <- relevel(data$type, "msm") #another way of changing order of levels
data1 <- xtabs(count ~type+case+year,data=data)
ftable(data1)
```
<br>
To perform the analysis we used a function to divide the dataset by year to calculate whether there was a difference between cases of HIV for MSM and IDU by specific years. The p-values were:
```{r, include=FALSE, warning=FALSE,message=FALSE}
data <- read.csv(".\\propt3.csv")
data$case <- factor(data$case, levels(data$case)[c(2,1)]) 
#one way of changing order of levels
#levels(data$case)
data$type <- relevel(data$type, "msm") #another way of changing order of levels
#levels(data$type)

prop <- function(x) {
        data <- filter(data, year == x)
        ma <- with(data, xtabs(count~ type +case))
        ex <- chisq.test(ma)$expected
        ex <- as.numeric(ex)
        if (any(ex < 5)){
                fisher.test(ma)
        }
        else{
                chisq.test(ma)
        }
}
prop("2004")
prop("2005")
prop("2006")
prop("2007")
prop("2008")
prop("2009")
prop("2010")
prop("2011")
prop("2012")
prop("2013")
prop("2014")
```

```{r, echo=FALSE, warning=FALSE,message=FALSE, error=FALSE}
a <- prop("2004")
b <- prop("2005")
c <- prop("2006")
d <- prop("2007")
e <- prop("2008")
f <- prop("2009")
g <- prop("2010")
h <- prop("2011")
i <- prop("2012")
j <- prop("2013")
k <- prop("2014")
lista <- list(a,b,c,d,e,f,g,h,i,j,k)
vec <- vector()
p_value <- sapply(1:11, function(x){
        if (x == 5 | x == 8) {
                vec[x] <- lista[[x]][[1]]
        }
        else {
                vec[x] <- lista[[x]][[3]]
        }
})
p_value <- round(p_value, digits = 4)
Year <- 2004:2014
tabla <- cbind(Year,p_value)
kable(xtable(tabla), format= "pandoc", digits = 4, align = "l")
```


We can see that for years 2005, 2008, 2012 the results were not statistically significant (`r p_value[c(2,5,9)]` respectively).

